---
title: "Final Project"
output: html_notebook
---


<h1>DATA CLEANING:<h1>


```{r}
chocolate = read.csv("chocolate.csv")

```



```{r}
str(chocolate)

```



```{r}

names(chocolate)[1:9] <- c("Company_Maker", "Bar_Name","REF", "Review_Date", "Cocoa_Percentage", "Company_Location", "Rating", "Bean_Type", "Bean_Origin")

names(chocolate)

```


```{r}
chocolate$Cocoa_Percentage <- as.numeric(sub("%", "", chocolate$Cocoa_Percentage))/100
str(chocolate)

```


```{r}

colSums(is.na(chocolate))
```



<h1>DATA EXPLORATION:<h1>




```{r}
hist(chocolate$Rating)

```



```{r}
cor.test(Rating,chocolate$REF)
```


```{r}
cor.test(Rating,Review_Date)

```


```{r}
cor.test(Rating,Cocoa_Percentage)
```



```{r}

pairs(chocolate[c("Rating","REF","Review_Date","Cocoa_Percentage")])

```



```{r}

plot(Rating~Company_Location, col="red")

```


```{r}
plot(Rating~Bar_Name, col="blue")

```


```{r}
plot(Rating~Company_Maker, col="red")

```


```{r}

plot(Rating~Bean_Type, col="blue")

```


```{r}
plot(Rating~Bean_Origin, col="yellow")


```



```{r}
library(ggplot2)
library(dplyr)

```



```{r}

select_country = chocolate %>% group_by(Company_Location) %>% summarize(count=n()) %>% arrange(desc(count)) %>% top_n(10)

ggplot(select_country, aes(x = reorder(Company_Location, -count), y = count))+
geom_bar(stat = 'identity', width = 0.7, col = "black", fill = "blue") +
labs(x = "Country", y = "Count", title = "Top 10 countries with more chocolate companies")

```


```{r}

company_make = chocolate %>% group_by(Company_Maker) %>% summarize(count=n()) %>% arrange(desc(count)) %>% top_n(10)

ggplot(company_make, aes(x = reorder(Company_Maker, -count), y = count))+
geom_bar(stat = 'identity', width = 0.7, col = "black", fill = "Red") +
labs(x = "Makers", y = "Count", title = "Top 10 chocolate company Makers")


```


```{r}

Bean = chocolate %>% group_by(Bean_Type) %>% summarize(count=n()) %>% arrange(desc(count)) %>% top_n(5)

ggplot(Bean, aes(x = reorder(Bean_Type, -count), y = count))+
geom_bar(stat = 'identity', width = 0.7, col = "black", fill = "green") +
labs(x = "Beans", y = "Count", title = "Top most used beans")


```



```{r}
Rate = chocolate %>% group_by(Rating) %>% summarize(count=n()) %>% arrange(desc(count)) 

ggplot(Rate, aes(x = reorder(Rating, -count), y = count))+
geom_bar(stat = 'identity', width = 0.7, col = "black", fill = "violet") +
labs(x = "Rating", y = "Count", title = "Given Ratings")


```



```{r}
chocolate$Rating = log(chocolate$Rating)

```


<h1>Creating Predictive Models: <h1>



```{r}
chocolate_raw = read.csv("chocolate.csv")

```


#This is the RMSE i got before cleaning(RAW) the data(coverting the cocoa_percentage to number), after cleaning the data it gave me a better results

```{r warning=FALSE}
set.seed(1)

trainn = createDataPartition(chocolate_raw$Rating, p=0.8, list = FALSE)

chocolate_raw_train = chocolate_raw[trainn,]
chocolate_raw_test = chocolate_raw[-trainn,]


lm_raw <- train(Rating~., data = chocolate_raw_train, method = "lm", trControl = trainControl("cv", number = 10))

predictions <- predict(lm_raw,chocolate_raw_test)
RMSE(predictions, chocolate_raw_test$Rating)

```

#This is the RMSE i got after cleaning the data, which has much better results. 



```{r}
library(caret)

Intrain = createDataPartition(chocolate$Rating, p=0.8, list = FALSE)

chocolate_train = chocolate[Intrain,]
chocolate_test = chocolate[-Intrain,]

```

<h1>"Multiple Linear Regression" without scaling.<h1>



```{r warning=FALSE}
set.seed(1)

lm <- train(Rating~., data = chocolate_train, method = "lm", trControl = trainControl("cv", number = 10))

predictions <- predict(lm,chocolate_test)
RMSE(predictions, chocolate_test$Rating)


```

#Scaling the data for better performance excluding the Target variable.

```{r}


chocolate_factors = chocolate[c(1,2,5,6,8,9)]

chocolate_numericals = chocolate[c(-1,-2,-5,-6,-7,-8,-9)]

chocolate_target = chocolate[7]

chocolate_factors = lapply(chocolate_factors, as.numeric)

scal = cbind(chocolate_factors,chocolate_numericals)

scaled = scale(scal)

chocolaten = cbind(scaled,chocolate_target)



```


```{r}

intrain = createDataPartition(chocolaten$Rating, p=0.8, list = FALSE)

chocolaten_train = chocolaten[intrain,]
chocolaten_test = chocolaten[-intrain,]


```


<h1>"Multiple Linear Regression" with scaling.<h1>

#We see almost 10% of change in RMSE when compared to whithout scaling data. Looking forward to fit the models with the scaled data.

```{r}
set.seed(1)

lm <- train(Rating~., data = chocolaten_train, method = "lm", trControl = trainControl("cv", number = 10))

summary(lm)


predictions <- predict(lm,chocolaten_test)
lm_RMSE = RMSE(predictions, chocolaten_test$Rating)

lm_RMSE


```

<h1>"Bagging Tree"<h1>

```{r}

set.seed(1)
library(caret)
Bag = train(Rating~., data = chocolaten_train, method = "treebag", nbagg=100, trControl=trainControl("cv", number=10))
Bag

```



```{r}
set.seed(1)
predictions<- predict(Bag, chocolaten_test)
Bag_rmse <- RMSE(predictions, chocolate_test$Rating)
Bag_rmse

```

<h1>"Regression Tree"<h1>


```{r}

set.seed(1)
library(caret)
library(rpart)
Rtrees = train(Rating~., data = chocolaten_train, method = "rpart1SE", trControl=trainControl("cv", number=10))
Rtrees

```



```{r}

set.seed(1)
predictions<- predict(Rtrees, chocolaten_test)
Rtrees_RMSE <- RMSE(predictions, chocolate_test$Rating)
Rtrees_RMSE

```

```{r}
library(rpart.plot)
rpart.plot(Rtrees$finalModel, roundint = FALSE)
```

<h1>"Model Tree"<h1>

```{r}
set.seed(1)
library(RWeka)
library(partykit)

Mtrees = train(Rating~., data = chocolaten_train, method = "M5", trControl=trainControl("cv", number=10))
Mtrees



```


```{r}

plot(Mtrees)
```



```{r}

set.seed(1)
predictions<- predict(Mtrees, chocolaten_test)
Mtrees_RMSE <- RMSE(predictions, chocolate_test$Rating)
Mtrees_RMSE
```

<h1>"Stepwise (Forward) Linear Regression"<h1>


```{r}

set.seed(1)

F.stepmodel <- train(Rating~., data = chocolate_train, method = "leapForward", trControl = trainControl(method = "cv", number = 10), tuneGrid = data.frame(nvmax = 1:22))
print(F.stepmodel)

summary(F.stepmodel$finalModel)


```



```{r}
set.seed(1)
predictions <- predict(F.stepmodel,chocolate_test)
F.stepmodel_RMSE = RMSE(predictions, chocolate_test$Rating)

F.stepmodel_RMSE


```

<h1>"Stepwise (Backward) Linear Regression"<h1>


```{r}

set.seed(1)

B.stepmodel <- train(Rating~., data = chocolate_train, method = "leapBackward", trControl = trainControl(method = "cv", number = 10), tuneGrid = data.frame(nvmax = 1:22))
print(F.stepmodel)

summary(B.stepmodel$finalModel)

```


```{r}
set.seed(1)
predictions <- predict(B.stepmodel,chocolate_test)
B.stepmodel_RMSE = RMSE(predictions, chocolate_test$Rating)

B.stepmodel_RMSE

```
<h1>"Regularized Linear Regression"<h1>

<h1>"Lasso"<h1>
```{r}
set.seed(1)

lasso <- train(Rating ~ ., data = chocolaten_train, method = "glmnet", trControl = trainControl("cv", number = 10), tuneGrid = expand.grid(alpha = 1, lambda = 10 ^ seq(-3, 3, length = 100)))

coef(lasso$finalModel, lasso$bestTune$lambda)

```



```{r}
set.seed(1)
predictions <- predict(lasso,chocolaten_test)
lasso_RMSE = RMSE(predictions, chocolaten_test$Rating)

lasso_RMSE

```

<h1>"Ridge"<h1>

```{r}
set.seed(1)
Ridge <- train(Rating ~ ., data = chocolaten_train, method = "glmnet", trControl = trainControl("cv", number = 10), tuneGrid = expand.grid(alpha = 0, lambda = 10 ^ seq(-3, 3, length = 100)))

coef(Ridge$finalModel, Ridge$bestTune$lambda)


```


```{r}
set.seed(1)
predictions <- predict(Ridge,chocolaten_test)
Ridge_RMSE = RMSE(predictions, chocolaten_test$Rating)

Ridge_RMSE

```
<h1>"ENET"<h1>

```{r}
set.seed(1)
ENET <- train(Rating ~ ., data = chocolaten_train, method = "glmnet", trControl = trainControl("cv", number = 10), tuneGrid = expand.grid(alpha = seq(0,1, length = 10), lambda = 10 ^ seq(-3, 3, length = 100)))

ENET

```


```{r}
set.seed(1)
ENET_predictions <- predict(ENET,chocolaten_test)
ENET_RMSE = RMSE(ENET_predictions, chocolaten_test$Rating)

ENET_RMSE

```
<h1>"Principal Componenet Analysis"<h1>


```{r}
PCA = prcomp(chocolaten_train, scale = TRUE)
names(PCA)

```

```{r}
PCA$x
```

```{r}
summary(PCA)
```

<h1>"RandomForest"<h1>

```{r}
set.seed(1)
library(caret)
rf = train(Rating ~ ., data = chocolaten_train, trControl = trainControl("cv", number = 10), tuneGrid = expand.grid(mtry = c(2, 4, 8, 16)), method = "rf", importance = TRUE)
rf

```




```{r}
set.seed(1)
predictions <- predict(rf,chocolaten_test)
rf_RMSE = RMSE(predictions, chocolaten_test$Rating)
rf_RMSE

```
<h1>"RandomForest with PCA"<h1>


```{r}
set.seed(1)
library(caret)
rf_pca = train(Rating ~ ., data = chocolaten_train, trControl = trainControl("cv", number = 10), preProcess = "pca", tuneGrid = expand.grid(mtry = c(2, 4, 8, 16)), method = "rf", importance = TRUE)

```



```{r}

rf_pca
rf_pca$preProcess

```

```{r}
set.seed(1)
predictions <- predict(rf_pca,chocolaten_test)
rf_pca_RMSE = RMSE(predictions, chocolaten_test$Rating)
rf_pca_RMSE

```

<h1>"Gradient Boosted Tree"<h1>

```{r}
set.seed(1)
library(caret)
gbm = train(Rating ~ ., data = chocolaten_train, trControl = trainControl("cv", number = 10), method = "gbm", preProc="nzv")
gbm
```


```{r}
set.seed(1)
gbm_predictions = predict(gbm, chocolaten_test)
gbm_RMSE <- RMSE(gbm_predictions, chocolate_test$Rating)
gbm_RMSE

```

Partioning the train data to 90% train and 10% validation.

```{r}

library(caret)

train = createDataPartition(chocolate_train$Rating, p=0.9, list = FALSE)



value = chocolate_train[train,]
chocolate_validation = chocolate_train[-train,]
chocolate_train1 = value
```

<h1>Artificial Neural Networks<h1>



```{r}

chocolate_train_cat = chocolate_train1[c(1,2,6,8,9)] 

chocolate_train_num = chocolate_train1[c(3,4,5)]

chocolate_train_target = chocolate_train1[7]

chocolate_test_cat = chocolate_test[c(1,2,6,8,9)] 

chocolate_test_num = chocolate_test[c(3,4,5)]

chocolate_test_target = chocolate_test[7]


chocolate_val_cat = chocolate_validation[c(1,2,6,8,9)] 

chocolate_val_num = chocolate_validation[c(3,4,5)]

chocolate_val_target = chocolate_validation[7]

```





```{r}
chocolate_train_cat = lapply(chocolate_train_cat, as.numeric)

str(chocolate_train_cat)

chocolate_test_cat = lapply(chocolate_test_cat, as.numeric)

str(chocolate_test_cat)

chocolate_val_cat = lapply(chocolate_val_cat, as.numeric)

str(chocolate_val_cat)

```



```{r}

chocolate_train_scale = cbind(chocolate_train_cat,chocolate_train_num)

chocolate_test_scale = cbind(chocolate_test_cat,chocolate_test_num)

chocolate_val_scale = cbind(chocolate_val_cat,chocolate_val_num)

```


```{r}

chocolate_train_scale=scale(chocolate_train_scale)

col_means_train <- attr(chocolate_train_scale, "scaled:center")
col_stddevs_train <- attr(chocolate_train_scale, "scaled:scale")


chocolate_test_scale = scale(chocolate_test_scale, center = col_means_train, scale = col_stddevs_train)

chocolate_val_scale <- scale(chocolate_val_scale, center = col_means_train, scale = col_stddevs_train)


chocolate_train_final = cbind(chocolate_train_scale,chocolate_train_target)

chocolate_test_final = cbind(chocolate_test_scale,chocolate_test_target)

chocolate_val_final = cbind(chocolate_val_scale,chocolate_val_target)



```



```{r}
train = chocolate_train_final

test = chocolate_test_final

val = chocolate_val_final



chocolate_train_final = train[-9]
chocolate_train_final_lables = train[9]


chocolate_test_final = test[-9]
chocolate_test_final_lables = test[9]

chocolate_val_final = val[-9]
chocolate_val_final_lables = val[9]


chocolate_train_final = as.matrix(chocolate_train_final)

chocolate_train_final_lables = as.matrix(chocolate_train_final_lables)

chocolate_test_final = as.matrix(chocolate_test_final)

chocolate_test_final_lables = as.matrix(chocolate_test_final_lables)

chocolate_val_final = as.matrix(chocolate_val_final)

chocolate_val_final_lables = as.matrix(chocolate_val_final_lables)

```



```{r}


set.seed(1)
library(keras)
library(tfruns)

runs <- tuning_run("chocolate.R",
flags = list(
nodes1 = c(64, 128),
nodes2 = c(128,392),
learning_rate = c(0.01, 0.05, 0.001, 0.0001),
batch_size=c(100,200,500,1000),
epochs=c(30,50,100),
activation=c("relu","sigmoid","tanh")
),
sample = 0.02
)


```



```{r}
print(runs)
```



```{r}
view_run(runs$run_dir[4])

```



```{r}

str(chocolate_train)

```


```{r}
train_factors = chocolate_train[c(1,2,6,8,9)]

train_numericals = chocolate_train[c(3,4,5)]

train_Rating = chocolate_train[7]

train_factors = lapply(train_factors, as.numeric)

train_scale = cbind(train_factors,train_numericals)


choco_train_scale=scale(train_scale)

choco_t = cbind(choco_train_scale,train_Rating)

val = choco_t

chocolate_t = val[-9]

chocolate_t_lables = val[9] 

chocolate_t = as.matrix(chocolate_t)

chocolate_t_lables = as.matrix(chocolate_t_lables)

```





```{r}
set.seed(111)

model =keras_model_sequential()

model %>%
  layer_dense(units = 64, activation = "sigmoid", input_shape = dim(chocolate_t)[2]) %>% layer_dropout(0.5) %>% 
  layer_dense(units = 128, activation = "sigmoid") %>% layer_dropout(0.5) %>%
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_adam(lr=0.05),
  loss = 'mse')

model %>% fit(
  chocolate_t, chocolate_t_lables, epochs = 50
  , batch_size= 200,
  validation_data=list(chocolate_val_final, chocolate_val_final_lables ))



predictions=model %>% predict(chocolate_test_final)

 rmse= function(x,y){
return((mean((x - y)^2))^0.5)
}
 ANN_RMSE = rmse(predictions,chocolate_test_final_lables)
 ANN_RMSE

```

<h1>RMSE COMPARISION<h1>


```{r}

RMSE_table <- matrix(c(lm_RMSE, Bag_rmse, Rtrees_RMSE, Mtrees_RMSE, F.stepmodel_RMSE, B.stepmodel_RMSE, lasso_RMSE, Ridge_RMSE, ENET_RMSE, rf_RMSE, rf_pca_RMSE, gbm_RMSE, ANN_RMSE), ncol = 13)

colnames(RMSE_table) <- c("Liner Regression", "Bagging Tree", "Regression Tress", "Model Trees", "Forward LRegression", "Backward LRegression", "Lasso", "Ridge", "ENET", "Randomforest", "Randomforestpca", "Gboost", "ANN")

rownames(RMSE_table) <- "RMSE"
RMSE_table <- as.table(RMSE_table)
RMSE_table

```


